---
title: "ESM 206 - Lab 5"
author: "Kirsten White"
date: "10/25/2021"
output: 
  html_document: 
    theme: lumen
    toc: yes
    code_folding: hide
---

### Learn cool customization 
- Wheel beside knit, output options (at the bottom), edit general, figures, advanced; add a theme (example: cerulean), table of contents (determined by hierarchy indicated by # signs), etc
- Code folding - gives user the option of seeing your code only if they want to. Do this by: adding code_folding: hide in the yaml 

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Add *.html in the .gitignore folder to ignore everything in html

# Attach Packages

library(tidyverse)
library(effsize) #install "effsize" package in the console 
```

## Part 1: Confidence Intervals

In this section, we'll use the `t.test()` to calculate confidence intervals 

### Example 1: 

Create a single mock sample containing a number of value, then learn to use `t.test()    to return confidence intervals, and learn about in-line code referencing so our text becomes more reproducible, too. 

```{r}
# Create a mock sample with a vector of 8 number values

otter_lengths <- c(30, 45, 28, 31, 34, 46, 52, 29)

# How do we get a confidence interval of these samples? Find confidence interval by using `t.test()`

# Returns statistical outcomes; most importantly a confidence interval (of 95%) - meaning that, if we took multiple samples and calculated a confidence interval for each, we can expect 95% of those intervals to contain the true population mean; this 29 to 44 will likely contain the population mean 
otter_ci <- t.test(otter_lengths)


# Find standard deviation 
otter_sd <- sd(otter_lengths)
```

Mean otter length is 36.9 inches +- ($/pm$) 9.32 inches (standard deviation = 9.32, n = 8) with a 95% confidence interval of [29.1 to 44.7] inches. (*** need to include sample size, standard deviation***)

*** ---- This is not reproducible by manually typing in these values. We want to build in ways to insert outputs of our code in our description. USE in-line referencing to outputs of code & store values as something & pulling the pieces by name. Find the documentation of these functions by searching ?t.test (example) in the console and reading the value section listed in the help area (example: statistic is the value of the t-statistic; conf.int is the value of the confidence interval--- recall it with object_name$value_name;  example: otter_ci$p.value or otter_ci$conf.int or otter_ci$statistic ****

#### Now, using in-line referencing

Mean otter length is (use one backtick + r to input mini code chunk) `r round(otter_ci$estimate, 1)` inches (to show value of the estimated mean and round it to one decimal place) 

